OptiFit: a fast method for fitting amplicon sequences to existing OTUs

### 2021-03-16 paper/paper.Rmd
Fix ref-db methods paragraph in results

- **TODO: current method for open/closed is vsearch against greengenes.**
- **TODO**: 2 categories of clustering: _de novo_ and reference based. separate paragraphs. describe opticlust first in de novo paragraph. 2nd paragraph: ref methods are good cause they're fast and don't use much ram. dependent on order of db. people use greengenes, which are rep seqs from 3% otus from full length.
- reader should know what opticlust is, closed & open ref clustering is, strengths & weakness are of each. then we solve these problems.
- **TODO:** note that greengenes is defunct now?!

@@

**TODO: make sure this info is baked into relevant sections, then remove.**
For each dataset repeated with 100 random seeds, we generated OTUs with OptiFit
using both strategies.
To compare to existing software, we also clustered OTUs _de novo_ using
OptiClust and VSEARCH, and with VSEARCH in reference-based mode against the
Greengenes database.

@@

**TODO: use word "map" for what vsearch does, "fit" for what optifit does.**

**TODO: ref clustering method paragraph here.**
We assigned blah ref dbs to 3% otus with opticlust. _de novo_ clustering the Greengenes
database, SILVA non-redundant database, or the Ribosomal Database Project (RDP)
[@desantis_greengenes_2006; @quast_silva_2013; @cole_ribosomal_2014].
All clustering was performed at a distance threshold of 0.03 and OTU quality was
evaluated using the MCC as described previously [@westcott_opticlust_2017].
To evaluate reference-based clustering with independent databases, we fit each
dataset to reference OTUs generated by...
repeated 100 times.

@@

We also tested three methods for selecting the fraction of sequences to be used
as the reference; a simple random sample, weighting sequences by relative
abundance, and weighting by similarity to other sequences in the dataset.
OTU quality was similar with the simple and abundance-weighted sampling (median
MCCs `r mcc_fit_split_simple` and `r mcc_fit_split_abun` respectively), but worse for similarity-weighted
sampling with a median MCC of `r mcc_fit_split_dist`.
In closed reference mode, the fraction of query sequences that can be fit to the
reference OTUs decreases as the reference fraction increases from 
`r frac_fit_split_0.1`% of query sequences fit with `10`% of the dataset as the
reference, to `r frac_fit_split_0.8`% of query sequences fit with `80`% of the
dataset as the reference.

>>>>

- **TODO: current method for open/closed is vsearch against greengenes.**
- **TODO: use word "map" for what vsearch does, "fit" for what optifit does.**
- **TODO**: 2 categories of clustering: _de novo_ and reference based. separate paragraphs. describe opticlust first in de novo paragraph. 2nd paragraph: ref methods are good cause they're fast and don't use much ram. dependent on order of db. people use greengenes, which are rep seqs from 3% otus from full length.
- reader should know what opticlust is, closed & open ref clustering is, strengths & weakness are of each. then we solve these problems.
- **TODO:** note that greengenes is defunct now?!

@@

@@

We performed _de novo_ clustering with OptiClust at distance threshold of 4%
using the Greengenes database, SILVA non-redundant database, or the Ribosomal
Database Project (RDP) [@desantis_greengenes_2006; @quast_silva_2013;
@cole_ribosomal_2014].
We then performed refernce clustering with OptiFit, fitting each dataset to each 
set of database OTUs repeated 100 times.
We measured the algorithm run time and evaluated OTU quality using the MCC as
described previously [@westcott_opticlust_2017].
To compare to existing software, we also clustered each dataset into _de novo_
OTUs  using OptiClust and VSEARCH, and with VSEARCH in reference-based mode
against the Greengenes database.

@@

We also tested three methods for selecting the fraction of sequences to be used
as the reference; a simple random sample, weighting sequences by relative
abundance, and weighting by similarity to other sequences in the dataset.
OTU quality was similar with the simple and abundance-weighted sampling (median
MCCs `r mcc_fit_split_simple` and `r mcc_fit_split_abun` respectively), but
worse for similarity-weighted sampling with a median MCC of `r mcc_fit_split_dist`.
In closed reference mode, the fraction of query sequences that can be fit to the
reference OTUs increases as the reference fraction increases; from 
`r frac_fit_split_0.1`% of query sequences fit when using 10% of the dataset as 
the reference, to `r frac_fit_split_0.8`% of query sequences fit when using 80% 
of the dataset as the reference.
